---
title: "Regression"
output: html_notebook
author: CS 4375.004 - Sunni Magan
date: 03/07/2023
---
## Introduction

This notebook explores the fuel consumption of cars from the years 2000-2022. 

Source: [https://www.kaggle.com/datasets/ahmettyilmazz/fuel-consumption](https://www.kaggle.com/datasets/ahmettyilmazz/fuel-consumption)

### About The Data

#### Model

- 4WD/4X4 = Four-wheel drive
- AWD = All-wheel drive
- CNG = Compressed natural gas
- FFV = Flexible-fuel vehicle
- NGV = Natural gas vehicle
- \# = High output engine that provides more power than the standard engine of the same size

#### Transmission

- A = Automatic
- AM = Automated manual
- AS = Automatic with select shift
- AV = Continuously variable
- M = Manual
- 3 - 10 = Number of gears

#### Fuel Type
- X = Regular gasoline
- Z = Premium gasoline
- D = Diesel
- E = Ethanol (E85)
- N = Natural Gas

#### Fuel Consumption
City and highway fuel consumption ratings are shown in litres per 100 kilometres (L/100 km) - combined rating (55% city, 45% hwy) is shown in L/100 km and in miles per imperial gallon (mpg)

CO2 Emissions (g/km),Estimated tailpipe carbon dioxide emissions (in grams per kilometre) are based on fuel type and the combined fuel consumption rating.

## Data Cleaning
Multiple factors in the database do not provide uniform data.
```{r}
df <- read.csv('Fuel_Consumption_2000-2022.csv')
```


```{r}
df$MAKE <- tolower(df$MAKE)
df$MAKE <- as.factor(df$MAKE)
```
The MAKE column provides car makes, however the capitalization is not consistant.

```{r}
df$VEHICLE.CLASS <- tolower(df$VEHICLE.CLASS)
df$VEHICLE.CLASS <- gsub(":", " -", df$VEHICLE.CLASS)
df$VEHICLE.CLASS <- as.factor(df$VEHICLE.CLASS)
```
The VEHICLE.CLASS column has a similar capitalization issue as MAKE. However, it also has differing punctuation.

```{r}
df$TRANSMISSION <- as.factor((df$TRANSMISSION))
df$FUEL <- as.factor(df$FUEL)
df$MODEL <- as.factor(df$MODEL)
```
The TRANSMISSION, FUEL, and MODEL columns are all uniform and simply need to be converted to a factor

## Train/Test Split

```{r}
set.seed(1234)
i <- sample(1:nrow(df), 0.8*nrow(df), replace=FALSE)
train <- df[i,]
test <- df[-i,]
```

## Data Exploration
```{r}
summary(df)
```